\plaintitle

### 2016-09-21 drafts/labor/piecework_lit.tex
added a ref shortname
piecework checkin

@@

\subsection{What was piecework?}
\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While ``piecework'' has had multiple definitions over time,
we can trace a constellation of characteristics that recur throughout the literature.
We will follow this history of research, collecting
descriptions,
examples, and
provided definitions of piecework, trying
to trace the outline of a working understanding of
\textit{what piecework is}.

@@

\topic{It's worth acknowledging that
``this distinction [between piece--rates and time--rates] was not completely clear--cut''
\cite{hart2013rise}.}
Indeed, we see work that adopts
piece--rate compensation in some aspects and
time--rate compensation in others.
The ``Rowan premium system'',
which essentially paid workers
a base rate for time plus
(the potential for) an additional pay dependent on output,
was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms, which
muddies the waters for us later as we attempt to categorize cases of piecework
\cite{rowan1901premium}.
As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
regardless of the worker's productive output
\textit{as well as} an additional compensation tied to performance,
workers under this regime were in some senses ``task--labourers'',
and in other senses
(more conventional)
``day--labourers''.
% \ali{I want to come back to this later when we talk about Uber \& Lyft, and
%      how they increasingly offer to guarantee drivers an hourly rate
%      \cite[see][]{uberHourly,lyftHourly}.
%      That's why I'm nitpicking about this muddied system --- to avoid someone saying
%      ``it doesn't seem like piecework to me anymore if there's an hourly wage floor''}

\topic{It may be worth thinking about piecework through the lens of its \textit{emergent} properties to help understand it.}
Returning to
\citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piecework
% \msb{inconsistent use of ``piece work'' vs. ``piecework'' in the non-quoted text. Which do you want?}
% \ali{Okay here's the problem: I'm happy to use ``piecework'' and that's the word that exists in the dictionary, but
% at the time that \citeauthor{hughRaynbirdTaskWork} was writing, it was ``piece work'' and I don't know if
% this is a trivial enough thing that I can just quietly concatenate his/their quotations so it looks like they say ``piecework''.
% I think at this point I was just calling it ``piece work'' because at this point in history people were calling it that.
% I've just changed everything out of quotation marks to ``piecework[er[s]]'', and
% left everything in quotation marks as ``piece work''.}
crop up; he points out that\dots
``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion\dots
he knows that all depends on his own diligence and perseverance\dots~[and]
so long as he performs his work to the satisfaction of his master,
he is not under that control to which the day--labourer is always subject.''
\citeauthor{hughRaynbirdTaskWork} (and others, as we will see)
highlight the freedom from control that ``task--labourers'' enjoy
\cite{hughRaynbirdTaskWork,rowan1901premium}.

@@

In summary, piecework:
\begin{enumerate}
  \item paid workers for quantity of work done, rather than time done, but occasionally mixed the two payment models;
  \item afforded workers freedom in when and how much to work; and
  \item structured tasks such that people who didn't have the training to engage in the traditional labor force could still participate.
\end{enumerate}
\ali{woah i thought this was a comment; is it okay as is?}

\ali{REMOVED A BUNCH OF STUFF FROM HERE ONWARD}

\subsection{The Historical Arc of Piecework}\label{sec:pieceworkArc}
\topic{Piecework's history traces back further perhaps than most would expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
to calculate equations in order
to better--predict the trajectories of various celestial bodies in the \nth{19} century
\cite{grier2013computers}.
While this approach didn't become the same economic powerhouse as later examples would,
Airy \msb{This is the first time Airy comes up. Who are they?} and others arguably found the kernel of insight that we pursue throughout this discussion:
determining the extent to which work can be decomposed, and
finding the limits of complexity of that decomposed work.
That is, Airy found that he could train youths in elementary mathematics
to complete the majority of the calculations he would otherwise have had to solve on his own,
and that the greater body of work could ultimately be completed sooner
if he arranged his work appropriately.
\msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
What I got out of it is that a bunch of people did piecework, but
I don't know why or in fact why these are different than the sources we cite earlier.
Can you hone it?}
\ali{I wanted that paragraph to be about the rising popularity and application of piecework
(especially as it approached its ``heyday'' \cite{hart2013rise}),
coming from humble beginnings as it found its footing.
Given that intent\dots\\
Should I \textbf{refactor} or \textbf{rewrite}?}

\ali{\sout{I'm thinking of dropping most of the last paragraph and merging it with the next one. Thoughts?}
I moved it to \nameref{sec:complexity}}

\ali{??? Why is this here? Can it be moved?}
\topic{As increasing attention revealed problems in piecework as it related to workers,
workers themselves began to speak out about their frustration with this new regime.}
It began, arguably, with \citeauthor{riisOtherSideLives}'s photo--documentary work,
but this led to industry organizations representing
railway workers, mechanical engineers, and others contributing their myriad perspectives
\cite{american1921problem,richards1904anything,riisOtherSideLives}.
Nevertheless, piecework continued to permeate low--skilled labor.
% \ali{How does this tie in with the rest of the narrative? Can I move this around? Does it break everything to take it out of chronological order?}

\ali{???}
\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance;
\item the shifting of these industries (manufacturing, clothing, etc\dots)
to other countries;
\item the quality of ``multidimensional'' work becoming too difficult to evaluate.
\end{inlinelist}
\cite{hart2013rise}.

>>>>

\ali{Maybe something here about how this isn't comprehensive, but it gives you enough to understand the
era about which we're talking, the very high--level things that were going on, etc\dots~so that
the later discussions actually make sense?}

@@

\subsection{What is piecework?}
\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While ``piecework'' has had multiple definitions over time,
we can trace a constellation of characteristics that recur throughout the literature.
We will follow this history of research, collecting
descriptions,
examples, and
provided definitions of piecework, trying
to trace the outline of a working understanding of
\textit{what piecework is}.

@@

\topic{It's worth acknowledging that
``this distinction [between piece--rates and time--rates] was not completely clear--cut''
\cite{hart2013rise}.}
Indeed, we see work that adopts
piece--rate compensation in some aspects and
time--rate compensation in others.
The ``Rowan premium system'',
which essentially paid workers
a base rate for time plus
(the potential for) an additional pay dependent on output,
was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms, which
muddies the waters for us later as we attempt to categorize cases of piecework
\cite{rowan1901premium}.
As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
regardless of the worker's productive output
\textit{as well as} an additional compensation tied to performance,
workers under this regime were in some senses ``task--labourers'',
and in other senses
(more conventional)
``day--labourers''.

\topic{It may be worth thinking about piecework through the lens of its \textit{emergent} properties to help understand it.}
Returning to
\citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piecework
crop up; he points out that\dots
``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion\dots
he knows that all depends on his own diligence and perseverance\dots~[and]
so long as he performs his work to the satisfaction of his master,
he is not under that control to which the day--labourer is always subject.''
\citeauthor{hughRaynbirdTaskWork} (and others, as we will see)
highlight the freedom from control that ``task--labourers'' enjoy
\cite{hughRaynbirdTaskWork,rowan1901premium}.

@@

In summary, piecework:
\begin{enumerate}
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models;
  \item afforded workers freedom in when and how much to work; and
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{enumerate}
% \ali{woah i thought this was a comment; is it okay as is?}

% \ali{REMOVED A BUNCH OF STUFF FROM HERE ONWARD}

\subsection{A Primer of Piecework}\label{sec:pieceworkPrimer} % yay rhyming?
\topic{In this section we will offer a brief overview of the history of piecework;
this should not be mistaken for a comprehensive or especially rigorous background.}
Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
In other words, framing piecework in the contexts of
the early days of the industrial revolution,
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
While the previous section provided a \textit{definition} of piecework,
this section attempts to shine a light on the \textit{zeitgeist} of piecework.

\topic{Piecework's history traces back further perhaps than most would expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
to calculate equations in order
to better--predict the trajectories of various celestial bodies in the \nth{19} century
\cite{grier2013computers}.
George Airy was perhaps the first to rigorously apply piecework--style decomposition of tasks to work;
by breaking complex calculations into constituent parts, and
training young men to solve simple algebraic problems,
Airy could distribute work to many more people than could otherwise complete the full calculations.

% While this approach didn't become the same economic powerhouse as later examples would,
% Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
% determining the extent to which work can be decomposed, and
% finding the limits of complexity of that decomposed work.
% Airy found that he could train youths in elementary mathematics
% to complete the majority of the calculations he would otherwise have had to solve on his own,
% and that the greater body of work could ultimately be completed sooner
% if he arranged his work appropriately.
% \msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
% What I got out of it is that a bunch of people did piecework, but
% I don't know why or in fact why these are different than the sources we cite earlier.
% Can you hone it?}
% \ali{I wanted that paragraph to be about the rising popularity and application of piecework
% (especially as it approached its ``heyday'' \cite{hart2013rise}),
% coming from humble beginnings as it found its footing.
% Given that intent\dots\\
% Should I \textbf{refactor} or \textbf{rewrite}?}

% \ali{\sout{I'm thinking of
% dropping most of the last paragraph and merging it with the next one. Thoughts?}
% I moved it to \nameref{sec:complexity}}

\ali{Where do I want to take this? Where are we now?
\begin{enumerate}
  \item Piecework started in the \nth{19} century (above). [DONE]
  \item Piecework went on to be a pretty cool thing in textiles and farming at the turn of the \nth{20} century. [DONE?]
  \item Piecework went gangbusters through the first half of the century, especially through the Second World War. []
  \item Piecework died pretty quickly (below). [DONE]
\end{enumerate}
Sound good?}

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took off on in farm work \cite{hughRaynbirdTaskWork},
in textiles \cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads \cite{Brown01041990}, and 
elsewhere in manufacturing \cite{10.2307/3827491}.
This proliferation of manual piecework led to discussion surrounding how best to manage these workers
\cite{norton1900textile,clark1908cotton},
which seemed to lead to frustration among workers over their poor working and living conditions
(famously documented by
% \ali{??? Why is this here? Can it be moved?}
% {As increasing attention revealed problems in piecework as it related to workers,
% workers themselves began to speak out about their frustration with this new regime.}
\citeauthor{riisOtherSideLives})
\cite{riisOtherSideLives}.
This led to industry organizations representing
railway workers, mechanical engineers, and others contributing their myriad perspectives
\cite{american1921problem,richards1904anything}.
% Nevertheless, piecework continued to permeate low--skilled labor.
% \ali{How does this tie in with the rest of the narrative? Can I move this around? Does it break everything to take it out of chronological order?}

\topic{Fordism and scientific management thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.}
\citeauthor{hart2013rise} point out that the Second World War,
which called millions of Americans to military service,
necessitated the rapid training and employment of
a labor pool that hadn't historically been utilized in industrial labor: women
\cite{hart2013rise}.
% With little time and virtually none of the training that was conventionally expected of metalworkers and other industrial work
% (e.g. years--long apprenticeships),
% tasks had to be redesigned and workers trained for these dramatically unorthodox task designs
\ali{I'd like to bring the Rosie the Riveter anecdote in here but it's not feeling natural\dots}
% \cite{honey1985creating,davies2014origins}.

\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance;
\item the shifting of these industries (manufacturing, clothing, etc\dots)
to other countries; and
\item the quality of ``multidimensional'' work becoming too difficult to evaluate
\end{inlinelist}
\cite{hart2013rise}.
